Nail: A practical interface generator for data formats

### 2014-12-03 paper/login/design.tex
pass through 2.1
one \XXX where i have no idea what the sentence meant

\section{Design By Example}
In this section, we will explain how to handle basic data formats in Nail, how to handle
redundancies in the format with dependent fields and how Nail parsers can be extended with
transformations.

As a motivating example, we will parse DNS packets, as defined in RFC 1035. Each DNS packet consists of a
header, a set of question records, and a set of answer records. Domain
names in both queries and answers are encoded as a sequence of labels,
terminated by a zero byte.  Labels are Pascal-style strings, consisting
of a length field followed by that many bytes comprising the label.
\label{s:design}
\subsection{Basic Data formats}

\input{fig-dnsfull}
Let's step through a simplified Nail grammar for DNS
packets, shown in  Figure~\ref{fig:dns-full}.  For this grammar, Nail produces the type declarations
shown in Figure~\ref{fig:dns-full-struct}, and the parser and
generator functions shown in Figure~\ref{fig:dns-full-api}.
Nail grammars are reuseable between applications and we will use this grammar to implement
both a DNS server and client, which previously would have had two separate hand-written parsers,
leading to bugs such as  the Android Master Key.

A Nail grammar file consists of rule definitions, for example l. 1-20,  which assign a name
(\texttt{dnspacket}) to a grammar production (2-20). If you are not familiar with
other parsers, you can imagine rules as C type declarations on steroids
(although our syntax is inspired by Go).

Just like C supports various constructs to build up types, such as structures and unions from pointers and
elemental types, Nail supports various \emph{combinators} to represent features of a file or protocol. We will present the features we used in implementing
DNS. A more complete reference can be found in~\ref{fig:syntax}, with detailed rationale in~\cite{bangert:nail-osdi14}.

\paragraph{Integers and Constraints}.
Because Nail is designed to cope with binary formats,
it handles not only common integer types (such as \texttt{uint16} on l. 16), but bit fields of any
length, such as \texttt{uint1}. These integers are exposed to the programmer as an appropriately
sized machine integer, e.g. \texttt{uint8\_t}. Nail also supports constraints on integer values,
limiting the values to either a range (l. 23, \texttt{|1..16}), which can optionally be half-open,
or a fixed set (l. 24, \texttt{|[1,255]}). Both types of constraint can be combined, e.g. \texttt{|
  [1..16,255]}. Constant values are also supported, e.g. l. 10: \texttt{uint3=0} represents
three reserved bits that must be 0. Because constant values carry no information, they are not
represented in the data type.

@@

Nail supports various forms of arrays. Line 40 shows how to parse a domain in a DNS packet with
\cc{many}, which keeps repeating the \cc{label} rule until it fails.  In the next section, we will
explain how to handle count fields, and our full paper describes how to handle various array
representations (such as delimiters or non-empty arrays).

>>>>

\section{Design By Example}
In this section, we will explain how to handle basic data formats in Nail, how to handle
redundancies in the format with dependent fields, and how Nail parsers can be extended with
transformations.

As a motivating example, we will parse DNS packets, as defined in RFC 1035. Each DNS packet consists of a
header, a set of question records, and a set of answer records. Domain
names in both queries and answers are encoded as a sequence of labels,
terminated by a zero byte.  Labels are Pascal-style strings, consisting
of a length field followed by that many bytes comprising the label.
\label{s:design}
\subsection{Basic data formats}

\input{fig-dnsfull}
Let's step through a simplified Nail grammar for DNS
packets, shown in  Figure~\ref{fig:dns-full}.  For this grammar, Nail produces the type declarations
shown in Figure~\ref{fig:dns-full-struct}, and the parser and
generator functions shown in Figure~\ref{fig:dns-full-api}.
Nail grammars are reusable between applications and we will use this grammar to implement
both a DNS server and a DNS client, which previously would have had two separate hand-written parsers,
leading to bugs such as the Android ZIP file vulnerability.

A Nail grammar file consists of rule definitions, for example lines 1--20,  which assign a name
(\texttt{dnspacket}) to a grammar production (lines 2--20). If you are not familiar with
other parsers, you can imagine rules as C type declarations on steroids
(although our syntax is inspired by Go).

Just like C supports various constructs to build up types, such as structures and unions from pointers and
elemental types, Nail supports various \emph{combinators} to represent features of a file or protocol. We will present the features we used in implementing
DNS. A more complete reference can be found in Figure~\ref{fig:syntax}, with detailed rationale in~\cite{bangert:nail-osdi14}.

\paragraph{Integers and Constraints.}
Because Nail is designed to cope with binary formats,
it handles not only common integer types (such as \texttt{uint16} on line 16), but bit fields of any
length, such as \texttt{uint1}. These integers are exposed to the programmer as an appropriately
sized machine integer, e.g. \texttt{uint8\_t}. Nail also supports constraints on integer values,
limiting the values to either a range (line 23, \texttt{|1..16}), which can optionally be half-open,
or a fixed set (line 24, \texttt{|[1,255]}). Both types of constraint can be combined, e.g. \texttt{|
  [1..16,255]}. Constant values are also supported; for instance, on line 10, \texttt{uint3=0} represents
three reserved bits that must be 0. Because constant values carry no information, they are not
represented in the data type.

@@

Nail supports various forms of arrays. Line 40 shows how to parse a domain in a DNS packet with
\cc{many}, which keeps repeating the \cc{label} rule until it fails.  In the next section, we will
explain how to handle count fields, and our full paper~\cite{bangert:nail-osdi14} describes how to handle various array
representations (such as delimiters or non-empty arrays).

### 2014-12-03 paper/login/intro.tex
pass through 2.1
one \XXX where i have no idea what the sentence meant

  -- Edward Shelley on the Android Master Key
\end{quote}
\section{Introduction}
Binary file formats and network protocols are hard to parse safely: 
The libpng image decompression library had 24 remotely
exploitable vulnerabilities from 2007 to 2013 according to CVEDetails,
Adobe's PDF and Flash viewers have been notoriously plagued by input
processing vulnerabilities, and even the zlib compression library had
input processing vulnerabilities in the past.
Most of these attacks involve memory corruption - therefore, it is easy to assume that 
solving memory corruption will end all our woes when handling untrusted inputs.

However, as memory-safe languages and exploit mitigation  tricks are becoming more prevalent,
attackers are moving to a new class of attack - parser differentials. Many applications use
hand-written input processing code, which is often mixed with the rest of the application, e.g. by
passing a pointer to the raw input through the application. This (anti-)pattern makes it impossible
to figure out whether two implementations of the same format or protocol are identical, and input
handling code can't be easily reused between applications. As a result, different applications often
disagree in the corner cases of a protocol, which can have fatal security consequences. For example,
Android has two parsers for ZIP archives involved in securely installing applications. First, a Java
program checks the signatures of files contained within an app archive and then another tool
extracts them to the filesystem. Because the two ZIP parsers disagree in multiple places, attackers
can modify a valid file so that the verifier will see the original contents, but attacker-controlled
files will be extracted to the filesystem, bypassing Android's code signing. Similar issues showed
up on iOS \cite{geohot-evasion} and SSL \cite{DBLP:conf/fc/KaminskyPS10}.

@@

However, if you wanted to use existing tools to parse PDF
or ZIP, you would soon find that they cannot handle the complicated -- and therefore most
error-prone-- aspects of such formats-- so you'd still have to
hand-write the riskiest bits of code. For example, existing parser generators cannot conveniently represent size or offset fields and more complex features, such as data compression or checksums cannot be expressed at all.

Furthermore, some parser generators are cumbersome to use when parsing binary data for several
reasons. First, many parser generators don't directly data structures, but call semantic actions
that the you write to construct a data structure. Therefore,you must
describe the format up to three times -- in the grammar, the data structure and the semantic
actions. Second, most parser generators only address parsing inputs, so you'd have to manually
construct outputs. Some parser generators, such as 
Boost.Spirit  allow generating output, but require you to write another set of semantic actions.

We address these challenges with Nail, a new parser generator for binary formats.
First, Nail grammars describe not only a format, but also a data type to represent it within the program.
Therefore you don't have to write semantic actions and type declarations and you can no longer syntactic validation and semantic processing.
Second, Nail will also take care of generating output from the same data type, without requiring you
to write more risky code or think about the format again.

Third,  Nail introduces two abstractions, \emph{dependent fields} and
\emph{transformations}, to elegantly handle problematic structures,
such as offset fields or checksums.  Dependent fields capture fields
in a protocol whose value depends in some way on the value or layout
of other parts of the format; for example, offset or length fields,
which specify the position or length of another data structure, fall
in this category.  Transformations allow you to escape the
generated code to modify the raw data and interact with dependent fields
in a controlled manner.

>>>>

  -- Edward Shelley on the Android Master Key
\end{quote}
\section{Introduction}
Binary file formats and network protocols are hard to parse safely: 
The libpng image decompression library had 24 remotely
exploitable vulnerabilities from 2007 to 2013 according to CVEDetails,
Adobe's PDF and Flash viewers have been notoriously plagued by input
processing vulnerabilities, and even the zlib compression library had
input processing vulnerabilities in the past.
Most of these attacks involve memory corruption; therefore, it is easy to assume that 
solving memory corruption will end all our woes when handling untrusted inputs.

However, as memory-safe languages and exploit mitigation  tricks are becoming more prevalent,
attackers are moving to a new class of attack called {\em parser differentials}. Many applications use
hand-written input processing code, which is often mixed with the rest of the application, e.g. by
passing a pointer to the raw input through the application. This (anti-)pattern makes it impossible
to figure out whether two implementations of the same format or protocol are identical, and input
handling code can't be easily reused between applications. As a result, different applications often
disagree in the corner cases of a protocol, which can have fatal security consequences. For example,
Android has two parsers for ZIP archives involved in securely installing applications. First, a Java
program checks the signatures of files contained within an app archive and then another tool
extracts them to the filesystem. Because the two ZIP parsers disagree in multiple places, attackers
can modify a valid file so that the verifier will see the original contents, but attacker-controlled
files will be extracted to the filesystem, bypassing Android's code signing. Similar issues showed
up on iOS~\cite{geohot-evasion} and SSL~\cite{DBLP:conf/fc/KaminskyPS10}.

@@

However, if you wanted to use existing tools to parse PDF
or ZIP files, you would soon find that they cannot handle the complicated---and therefore most
error-prone---aspects of such formats, so you'd still have to
hand-write the riskiest bits of code. For example, existing parser generators cannot conveniently represent size or offset fields, and more complex features, such as data compression or checksums, cannot be expressed at all.

Furthermore, some parser generators are cumbersome to use when parsing binary data for several
reasons. First, many parser generators don't produce convenient data structures, and instead call
but call \emph{semantic actions} that the developer has to write to build up a data structure that
the rest of the program can use. Therefore, the developer must
describe the format up to three times: in the grammar, in the data structure, and in the semantic
actions. Second, most parser generators only address parsing inputs, so the developer would have to manually
construct outputs. Some parser generators, such as 
Boost.Spirit, allow generating output, but require the developer to write another set of semantic actions.

We address these challenges with Nail, a new parser generator for binary formats.

First, Nail grammars describe not only a format, but also a data type to represent the format within the program.
Therefore the developer doesn't have to write semantic actions and type declarations.
\XXX[what did this mean?  and can no longer syntactic validation and semantic processing.]

Second, Nail will also generate output from this data type, without requiring the developer
to write more risky code or giving the developer a chance to introduce ambiguity.

Third,  Nail introduces two abstractions, \emph{dependent fields} and
\emph{transformations}, to elegantly handle problematic structures,
such as offset fields or checksums.  Dependent fields capture fields
in a protocol whose value depends in some way on the value or layout
of other parts of the format; for example, offset or length fields,
which specify the position or length of another data structure, fall
in this category.  Transformations allow the developer to write plug-ins,
allowing programs to handle complicated structures, while
keeping Nail itself small, yet flexible.
